---
title: "florida_processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
# load packages
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(maps)
library(rgdal)
library(raster)
library(sdmpredictors)
library(here)

rasterOptions(maxmemory = 1e10)
```

This Rmarkdown:

- imports layers of interest
- reprojects, crops, and masks them to the area of interest 
- If shapefile, reclassifies to raster with 1 for fill and 0 for not then exports as raster
- exports them

Area of Interest bounding box
```{r aoi}
# # aoi shapefile
# aoi <- read_sf("../data/gulf_of_mexico_aoi/",
#                 layer = "gulf_of_mexico_aoi-polygon")

#projection
proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```

Base Raster
This is from the SABGOM model and cropped to the same extent as that data
```{r}
base <- raster(here("data", "FL_extent", "base_raster_FL.tif"))

aoi <- extent(base)
```


EEZ 
```{r}
# read in world eez
eez <- read_sf(here("data", "World_EEZ_v11_20191118"), layer = "eez_v11")

# filter to us eez and reproject
us_eez <- eez %>% 
  filter(TERRITORY1 == "United States") %>% 
  st_transform(proj)

# crop to Florida
eez_fl <- st_crop(us_eez, aoi) 
plot(eez_fl$geometry)

```


### GEBCO 

Depth Data
```{r}
# Get depth data of the gulf 
depth <- raster(here("data", "GEBCO_2020_28_Jul_2020_f2c031699f98", "gebco_2020_n34.32197570800781_s5.460205078125_w-101.04194641113281_e-69.25712585449219.tif"))

# reproject
depth_proj <- projectRaster(depth, crs = proj)

# crop depth to EEZ
depth_crop <- crop(depth, aoi)

plot(depth_crop)

```

###  Resample

```{r}
# Resample to SABGOM Cell Size (ATM Bio-Oracle Cell Size is the Standard
  
  # resample
  depth_resample_bo <- resample(depth_crop, base, method='ngb', progress='text')

# mask depth to EEZ
depth_mask <- mask(depth_resample_bo, eez_fl)

plot(depth_mask)
```

# Exclusion Layers

## Natural Resources

Marine Protected Areas

```{r}
# read in MPAs
mpa <- read_sf(here("data", "MPAI2017"), layer = "MPAI_2017")
# reproject
mpa_project <- mpa %>% 
  dplyr::select(Site_ID, Site_Name, Gov_Level, State) %>% 
  st_transform(proj) %>% 
  st_buffer(0)
# Crop
mpa_crop <- st_crop(mpa_project, aoi)
# Rasterizing Steps for MPA Layer
  # Create a Binary MPA Raster
  mpas_binary <- rasterize(mpa_crop, base, field = 0, background = 1) %>% 
    mask(eez_fl)
  
  # Plot Reclassified MPA Layer
  plot(mpas_binary)
  #freq(mpas_binary)
# Write a Raster for Suitability Analysis
writeRaster(mpas_binary, here("fl_output", "cropped_layers", "fl_mpa.tif"), overwrite = T)
```

## Travel lanes

Shipping lanes
```{r}
# read in shipping lanes
shipping <- read_sf(here("data", "shippinglanes"), layer = "shippinglanes")

# reproject 
ship_project <- shipping %>% 
  st_transform(proj) 

# crop
ship_crop <- ship_project %>% 
  st_crop(eez_fl)

# filter to only include shipping lanes and fairways
ship_filter <- ship_crop %>% 
  filter(THEMELAYER == "Shipping Fairways Lanes and Zones" | 
           THEMELAYER == "Traffic Separation Schemes" |
           THEMELAYER == "Traffic Separation Schemes/Traffic Lanes")

# rasterize & mask
ship_binary <- rasterize(ship_filter, base, field = 0, background = 1) %>% 
    mask(eez_fl)

plot(ship_binary)

# Write a Raster for Suitability Analysis
writeRaster(ship_binary, here("fl_output", "cropped_layers", "ship_binary_FL.tif"), overwrite = T)

```

## Infrastructure

Oil and gas platforms

```{r}
# read in oil and gas platforms 
og_platform <- read_sf(here("data", "OilandGasPlatforms", "OilandGasPlatforms.gdb"),
                       layer = "OilandGasPlatforms")

# add 500 meter buffer around points
og_buffer <- st_buffer(og_platform, 500)

# reproject and crop
og_crop <- og_buffer %>% 
  st_transform(proj) %>% 
  st_crop(aoi)

# rasterize & mask
og_binary <- rasterize(og_crop, base,
                                   field = 0, background = 1) %>% 
  mask(eez_fl)

# plot it
plot(og_binary)
map('world',fill=T,add=T,col='gray')

# Write a Raster for Suitability Analysis
writeRaster(og_binary, here("fl_output", "cropped_layers", "og_binary_hy.tif"), overwrite = T)
```

Submarine cable areas

```{r}
# read in submarine cables
sub_cable_area <- read_sf(dsn = here("data", "SubmarineCableAreas", "SubmarineCableAreas.gdb"),
                      layer = "SubmarineCableAreas" )

# reproject 
sub_cable_area_project <- sub_cable_area %>% 
  st_transform(proj) 

# crop
sub_cable_area_crop <- st_crop(sub_cable_area_project, aoi)

plot(sub_cable_area_crop)

# rasterize & mask
sub_cable_area_binary <- rasterize(sub_cable_area_crop, base,
                                   field = 0, background = 1) %>% 
  mask(eez_fl)

plot(sub_cable_area_binary)
map('world',fill=T,add=T,col='gray')

# Write a Raster for Suitability Analysis
writeRaster(sub_cable_area_binary, here("fl_output", "cropped_layers", "sub_cable_area_binary_fl.tif"), overwrite = T)

```

Pipeline areas
```{r}
# read in pipeline areas
pipeline_areas <- read_sf(dsn = here("data", "PipelineArea", "PipelineArea.gdb"),
                          layer = "PipelineArea")

# reproject
pipe_project <- st_transform(pipeline_areas, proj)

# crop
pipe_crop <- st_crop(pipe_project, aoi)

plot(pipe_crop$Shape)

# rasterize and mask
pipe_binary <- rasterize(pipe_crop, base,
                                   field = 0, background = 1) %>% 
  mask(eez_fl)

plot(pipe_binary)

# Write a Raster for Suitability Analysis
writeRaster(pipe_binary, here("fl_output", "cropped_layers", "pipe_binary_hy.tif"), overwrite = T)

```

## Military

Danger and restricted zones
```{r}

# read in sf

danger_zones <- st_read(dsn = here("data", "DangerZonesAndRestrictedAreas", "DangerZonesandRestrictedAreas.gdb"), 
                        layer = "DangerZonesandRestrictedAreas")

# reproject
danger_reproject <- st_transform(danger_zones, proj) 

# recast to multipolygon for cropping
danger_mp <- st_cast(danger_reproject, "MULTIPOLYGON")

# crop
danger_crop <- st_crop(danger_mp, aoi)

# rasterize and mask
danger_binary <- rasterize(danger_crop, base,
                                   field = 0, background = 1) %>% 
  mask(eez_fl)

plot(danger_binary)

# Write a Raster for Suitability Analysis
writeRaster(danger_binary, here("fl_output", "cropped_layers", "danger_binary_hy.tif"), overwrite = T)

```

All suitable areas
```{r}
suitable <- overlay(ship_binary, mpas_binary, sub_cable_area_binary, pipe_binary,
                    danger_binary, og_binary, 
                    fun = function(a, b, c, d, e, f){a*b*c*d*e*f})
plot(suitable)
map('world',fill=T,add=T,col='gray')

writeRaster(suitable, here("fl_output", "cropped_layers", "fl_exclusion.tif"), overwrite = TRUE)

res(suitable)
```