---
title: "data_exploration"
author: "Gabriel De La Rosa"
date: "7/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(raster)
library(sf)
library(here)
library(readxl)
library(fasterize)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(gstat)
```

### Set Projections and base rasters:
```{r}
# This is a WGS84 standard projection
proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

# This is a 0.2 degree resolution raster for the whole globe...
x <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=0.2, crs="+proj=longlat +datum=WGS84")

# Make a US basemap:

if(!file.exists(here("cropped_rasters", "US_mainland.shp"))){
  # Read in country data
  US <- ne_countries(type = "countries", country = "united states of america", scale = 50)
  # Convert it to an sf object
  us_sf <- st_as_sf(US)
  # Crop to the EEZ bounding box (returns only mainland)
  US_crop <- st_crop(us_sf, us_eez)
  # Save it for later use
  write_sf(US_crop, here("cropped_rasters", "US_mainland.shp"))
}

US_crop <- read_sf(here("cropped_rasters", "US_mainland.shp"))
```


```{r}
# This is a function to crop and mask a raster to the US EEZ.
mask_fx <- function(object){
  cropped <- raster::crop(object, us_eez)
  mask <- raster::mask(cropped, us_eez)
  return(mask)
}
```


### Create Area of Interest: US Mainland EEZ

``` {r}
if(!file.exists(here("cropped_rasters", "us_eez.shp"))){
  # Read in eez data. This is already in the proper projection.
  eez <- read_sf(here("data", "World_EEZ_v11_20191118"), layer = "eez_v11")
  
  # Filter it down
  us_eez <- eez %>% 
    filter(SOVEREIGN1 == "United States") %>% 
    filter(TERRITORY1 == "United States") %>% 
    filter(MRGID == "8456") %>% 
    dplyr::select(TERRITORY1)
  
  # Save just the US...
  sf::write_sf(us_eez, here("cropped_rasters", "us_eez.shp"))
}
  
# We can read in our processed files here to save time later:
us_eez <- read_sf(here("cropped_rasters"), layer = "us_eez")

# Convert this to an sp df...do I need this?
us_sp <- as(us_eez, 'Spatial')
```


### Eutrophication and hypoxia data from WOI
Read in eutrophication and hypoxia data:
This is point data for eutrophic and hypoxic regions in the US.
This is in WGS84 projection originally.

```{r}
if(!file.exists(here("cropped_rasters", "US_hypox_1km_rast.tif"))){
  #-------------
  # Read this in as a shapefile
  #-------------
  # Read Data
  eut_data<- read_xls(here("data", "wri_eutro_hypox", "wri_eutrophic_hypoxic_dataset_2011-03.xls"))
  # Convert it to spatial:
  eut_sf <- sf::st_as_sf(eut_data, coords = c("Long", "Lat"), crs = proj)
  # Crop to the US EEZ
  eut_US <- sf::st_crop(eut_sf, us_eez)
  # Write as cropped shapefile
  sf::write_sf(eut_US, here("cropped_rasters", "US_hypox.shp"))
  ##------------
  ## Then, convert it to a raster
  ##------------
  # First, get rid of the improved locations:
  eut_current <- eut_US %>% 
    filter(Classification %in% c("Hypoxic", "Eutrophic", "eutrophic")) %>% 
    dplyr::select(ID)
  # Put this into a 1 degree grid...
  eut_raster <- rasterize(eut_current, x, fun = 'count')
  # Crop it to the US EEZ...
  eut_us_crop <- raster::crop(eut_raster, us_eez)
  # Mask it to the US EEZ...
  eut_us_mask <- raster::mask(eut_us_crop, us_eez, snap = "out")
  # Save this raster
  writeRaster(eut_us_crop, here("cropped_rasters", "US_hypox_crop.tif"))
  writeRaster(eut_us_mask, here("cropped_rasters", "US_hypox_rast.tif"))
}

eut_us_crop <- raster(here("cropped_rasters", "US_hypox_crop.tif"))
eut_us_mask <- raster(here("cropped_rasters", "US_hypox_1km_rast.tif"))
plot(eut_us_crop)

```


### OHI Nutrient Data
These are the relative intensity of fertilizer pollution globally.
This is originally a mollewide projection.

```{r}
# This reads in the OHI raster, reprojects it into WGS84, then resamples it to 2degree resolution.

if(!file.exists(here("cropped_rasters", "OHI_nutrient_pollution_reproject.tif"))){
  # Read in data
  nut_raw <- raster(here("data", "ohi_2013", "nutrient_pollution_2013_raw.tif"))
  # Crop the raster to US bounding box
  nut_crop <- raster::crop(nut_raw, us_eez_moll)
  # Mask the raster to just the US EEZ
  nut_mask <- raster::mask(nut_crop, us_eez_moll, snap = "out")
  # Then, save the masked raster
  raster::writeRaster(nut_mask, here("cropped_rasters", "nut_mask.tif"))
  # Read this in again...
  nut_mask <- raster(here("cropped_rasters", "nut_mask.tif"))
  # Reproject the masked raster to x, our base raster with a 0.2 degree resolution
  nut_reproject <- projectRaster(nut_mask, crs = proj, progress = "text")
  # Save the reprojected layer
  writeRaster(nut_reproject, here("cropped_rasters", "OHI_nutrient_pollution_reproject.tif"))
}

nut_reproject <- raster(here("cropped_rasters", "OHI_nutrient_pollution_reproject.tif"))

# This resamples the raster to 2 degrees, using a bilinear method
if(!file.exists(here("cropped_rasters", "OHI_nut_resample_2dg.tif"))){
  nut_resample <- raster::resample(nut_reproject, x, method = "bilinear", progress = "text")
  nut_resample_mask <- mask_fx(nut_resample)
  writeRaster(nut_resample_mask, here("cropped_rasters", "OHI_nut_resample_2dg.tif"))
}

nut_resample_mask <- raster(here("cropped_rasters", "OHI_nut_resample_2dg.tif"))

plot(nut_resample_mask)
```


### Nitrate from Halley's paper.
This is originally in wgs84.
This is super low resolution (1 degree.)
These just include the mean of 0-10m measurements.

#### DANGER DANGER THIS LOOKS LIKE IT HAS MORE RESOLUTION THAN IT ACTUALLY DOES

```{r, eval = FALSE}
# Let's just use this...
if(!file.exists(here("cropped_rasters", "nitro_mean_2dg_2013.tif"))){
  # Read it in as a raster...
  fall_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Fall.tif"))
  # Crop and mask it to the US EEZ:
  fall_nut_mask <- mask_fx(fall_nut)
  # Repeat for spring, summer, fall
  # SPRING
  # Read it in as a raster...
  spring_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Spring.tif"))
  # Crop and mask it to the US EEZ:
  spring_nut_mask <- mask_fx(spring_nut)
  # SUMMER
  summer_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Summer.tif"))
  # Crop and mask it to the US EEZ:
  summer_nut_mask <- mask_fx(summer_nut)
  # WINTER
  winter_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Winter.tif"))
  # Crop and mask it to the US EEZ:
  winter_nut_mask <- mask_fx(winter_nut)
  
  # Stack these rasters:
  nitro_stack <- raster::stack(fall_nut_mask, spring_nut_mask, summer_nut_mask, winter_nut_mask)
   # Then, find the mean...
  nut_mean <- raster::calc(nitro_stack, fun = mean)
  # Then, resample down to the 0.2 degree resolution
  nitro_mean_resample <- raster::resample(nut_mean, x)
  # Then, recrop and mask:
  nitro_mean_resample_mask <- mask_fx(nitro_mean_resample)
  # Then, write raster
  writeRaster(nitro_mean_resample_mask, here("cropped_rasters", "nitro_mean_2dg_2013.tif"))
}

nitro_mean <- raster(here("cropped_rasters", "nitro_mean_2dg_2013.tif"))

plot(nitro_mean, nut_resample_mask)
```


### Phosphate from Halley's paper.
This is originally in wgs84.
This is super low resolution (1 degree.)
These just include the mean of 0-10m measurements.

#### DANGER DANGER THIS LOOKS LIKE IT HAS MORE RESOLUTION THAN IT ACTUALLY DOES

```{r, eval = FALSE}
# Let's just use this...
if(!file.exists(here("cropped_rasters", "phos_mean_2dg_2013.tif"))){
  # Read it in as a raster...
  fall_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Fall.tif"))
  # Crop and mask it to the US EEZ:
  fall_phos_mask <- mask_fx(fall_phos)
  # Repeat for spring, summer, fall
  # SPRING
  # Read it in as a raster...
  spring_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Spring.tif"))
  # Crop and mask it to the US EEZ:
  spring_phos_mask <- mask_fx(spring_phos)
  # SUMMER
  summer_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Summer.tif"))
  # Crop and mask it to the US EEZ:
  summer_phos_mask <- mask_fx(summer_phos)
  # WINTER
  winter_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Winter.tif"))
  # Crop and mask it to the US EEZ:
  winter_phos_mask <- mask_fx(winter_phos)
  
  # Stack these rasters:
  phos_stack <- raster::stack(fall_phos_mask, spring_phos_mask, summer_phos_mask, winter_phos_mask)
   # Then, find the mean...
  phos_mean <- raster::calc(phos_stack, fun = mean)
  # Then, resample down to the 0.2 degree resolution
  phos_mean_resample <- raster::resample(phos_mean, x, method = )
  # Then, recrop and mask:
  phos_mean_resample_mask <- mask_fx(phos_mean_resample)
  # Then, write raster
  writeRaster(phos_mean_resample_mask, here("cropped_rasters", "phos_mean_2dg_2013.tif"))
}

phos_mean <- raster(here("cropped_rasters", "phos_mean_2dg_2013.tif"))

plot(phos_mean)
plot(US_crop, add = TRUE, col = "grey80")
```

## Okay, there's missing data in these rasters, especially for shoreline values. I'm going to interpolate using an Inverse Distance Weighting function and methods/code from (OHI)[http://ohi-science.org/ohiprep_v2019/globalprep/prs_oa/v2017/create_oa_layer.html].

I need to check and see if this is legit...
```{r}
inter <- function(file){
  r <- file             # Load in the raster
  # grid <- raster::projectExtent(r, res = 0.2)
  xy <- data.frame(xyFromCell(r, 1:ncell(r))) #create a df of the grid
  v = getValues(r) # and a list of the values..
  tmpdf <- cbind(xy, v)%>%filter(!is.na(v)) # Bind them together
  mg <- gstat(id = "v", formula = v~1, locations = ~x+y, data=tmpdf,
            nmax=7, set=list(idp = 2)) # Create a model (power function = 2)
  z <- interpolate(r, mg, progress='text')
}

# Run this for both phosphate and nitrate
phos_inter <- inter(phos_mean)
nitro_inter <- inter(nitro_mean)

# Mask the values to the US EEZ: 
phos_int_mask <- mask_fx(phos_inter)
nitro_int_mask <- mask_fx(nitro_inter)

writeRaster(phos_int_mask, here("cropped_rasters", "phos_int_mask.tif"))
writeRaster(nitro_int_mask, here("cropped_rasters", "nitro_int_mask.tif"))
```
