---
title: "hotspot_processing"
author: "Gabriel De La Rosa"
date: "7/1/2020"
output: html_document
---

# This R Markdown processes files for the preliminary analysis presented early in the summer. 

## The layers processed here include:
 
* Eutrophication and hypoxia data from WOI
* OHI Nutrient Data - relative intensity of fertilizer pollution globally
* Nitrate & Phosphate from Halley's paper.
* Bio Oracle Nitrate & Phosphate
* Sewage Outfall Data

## The outputs from this document are all included in the cropped_rasters folder.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(raster)
library(sf)
library(here)
library(readxl)
library(fasterize)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(rgeos)
library(gstat)
library(doParallel)
library(rgis)
library(snow)


rasterOptions(maxmemory = 1e10, progress = "text")
```

### Set Projections and base rasters:
```{r}
# This is a WGS84 standard projection
proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

# This is a 0.2 degree resolution raster for the whole globe...
x <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res=0.2, crs= proj)
```


### Create Area of Interest: US  EEZ
``` {r}
if(!file.exists(here("cropped_rasters", "us_eez_ak_hi.shp"))){
  # Read in eez data. This is already in the proper projection.
  eez <- read_sf(here("data", "World_EEZ_v11_20191118"), layer = "eez_v11")
  
  # Filter it down
  us_eez <- eez %>% 
    filter(TERRITORY1 %in% c("United States")) %>%
    filter(MRGID == "8456") %>% 
    dplyr::select(TERRITORY1)
  
  ak_eez <- eez %>% 
    filter(TERRITORY1 %in% c("Alaska")) %>% 
    dplyr::select(TERRITORY1)
          
  hi_eez <- eez %>% 
    filter(TERRITORY1 %in% c("Hawaii")) %>% 
    dplyr::select(TERRITORY1)
  
  
  # Save just the US...
  sf::write_sf(us_eez, here("cropped_rasters", "us_eez.shp"))
  sf::write_sf(ak_eez, here("cropped_rasters", "ak_eez.shp"))
  sf::write_sf(hi_eez, here("cropped_rasters", "hi_eez.shp"))
}
  
# We can read in our processed files here to save time later:
us_eez <- read_sf(here("cropped_rasters"), layer = "us_eez")
ak_eez <- read_sf(here("cropped_rasters"), layer = "ak_eez")
hi_eez <- read_sf(here("cropped_rasters"), layer = "hi_eez")
```


```{r}
# Create a function to mask a raster to the US EEZ
mask_fx <- function(object, shp = us_eez){
  cropped <- raster::crop(object, shp)
  raster <- raster::rasterize(shp, cropped, getCover = TRUE)
  raster[raster == 0] <- NA
  mask <- rgis::fast_mask(cropped, raster)
  return(mask)
}
```

### US basemap and state lines:
```{r}
# Make a US basemap:

if(!file.exists(here("cropped_rasters", "US_mainland.shp"))){
  # Read in country data
  US <- ne_countries(type = "countries", country = "united states of america", scale = 50)
  # Convert it to an sf object
  us_sf <- st_as_sf(US)
  # Crop to the EEZ bounding box (returns only mainland)
  US_crop <- st_crop(us_sf, us_eez)
  # Save it for later use
  write_sf(US_crop, here("cropped_rasters", "US_mainland.shp"))
}

US_crop <- read_sf(here("cropped_rasters", "US_mainland.shp"))

# State Lines

if(!file.exists(here("cropped_rasters", "US_states.shp"))){
  # Read in country data
  US_state <- ne_states(country = "united states of america")
  # Convert it to an sf object
  us_state_sf <- st_as_sf(US_state)
  # Crop to the EEZ bounding box (returns only mainland)
  US_state_crop <- st_crop(us_state_sf, us_eez)
  # Save it for later use
  write_sf(US_state_crop, here("cropped_rasters", "US_states.shp"))
}
```


### Eutrophication and hypoxia data from WOI
Read in eutrophication and hypoxia data:
This is point data for eutrophic and hypoxic regions in the US.
This is in WGS84 projection originally.

```{r}
if(!file.exists(here("cropped_rasters", "US_hypox_rast.tif"))){
  #-------------
  # Read this in as a shapefile
  #-------------
  # Read Data
  eut_data<- read_xls(here("data", "wri_eutro_hypox", "wri_eutrophic_hypoxic_dataset_2011-03.xls"))
  # Convert it to spatial:
  eut_sf <- sf::st_as_sf(eut_data, coords = c("Long", "Lat"), crs = proj)
  # Crop to the US EEZ
  eut_US <- sf::st_crop(eut_sf, us_eez)
  # Write as cropped shapefile
  sf::write_sf(eut_US, here("cropped_rasters", "US_hypox.shp"))
  ##------------
  ## Then, convert it to a raster
  ##------------
  # reload in data:
  eut_US <- read_sf(here("cropped_rasters"), layer = "US_hypox")
  # First, get rid of the improved locations:
  eut_current <- eut_US %>% 
    filter(HypxCrr %in% c("Hypoxic", "Eutrophic", "eutrophic")) %>% 
    dplyr::select(ID)
  # Put this into a 0.2 degree grid...
  eut_raster <- rasterize(eut_current, x, fun = 'count')
  # Crop it to the US EEZ...
  eut_us_mask <- mask_fx(eut_raster)
  # Save this raster
  writeRaster(eut_us_mask, here("cropped_rasters", "US_hypox_rast.tif"), overwrite = TRUE, append = FALSE)
}



eut_us_mask <- raster(here("cropped_rasters", "US_hypox_rast.tif"))


plot(eut_us_mask)

```


### OHI Nutrient Data
These are the relative intensity of fertilizer pollution globally.
This is originally a mollewide projection.

```{r}
# This reads in the OHI raster, reprojects it into WGS84, then resamples it to 2degree resolution.

if(!file.exists(here("cropped_rasters", "OHI_nutrient_pollution_reproject.tif"))){
  # Read in data
  nut_raw <- raster(here("data", "ohi_2013", "nutrient_pollution_2013_raw.tif"))
  # Convert the EEZ projection to Mollewide...
  us_eez_moll <- st_transform(us_eez, crs(nut_raw))
  ak_eez_moll <- st_transform(ak_eez, crs(nut_raw))
  hi_eez_moll <- st_transform(ak_eez, crs(nut_raw))
  # Crop the raster to US bounding box
  nut_crop_us <- mask_fx(nut_raw, us_eez_moll) ## I finished this
  writeRaster(nut_crop_us, here("temp_rasters", "nut_mask_us.tif"))
  nut_crop_hi <- mask_fx(nut_raw, hi_eez_moll) #@ I finished this...
  writeRaster(nut_crop_hi, here("temp_rasters", "nut_mask_hi.tif"))
  
  ## ---------------
  ## ---------------
  #### I'm here
  ## ---------------
  ## ---------------
  beginCluster(detectCores() - 1)
  nut_crop_ak <- mask_fx(nut_raw, ak_eez_moll)
  endCluster()
  
  nut_reproj_us <- projectRaster(nut_crop_us, crs = proj, progress = "text")
  nut_reproj_hi <- projectRaster(nut_crop_hi, crs = proj, progress = "text")
  nut_reproj_ak <- projectRaster(nut_crop_ak, crs = proj, progress = "text")
  
  plot(us_eez_moll)
  
  # Then, save the masked raster
  raster::writeRaster(nut_mask, here("cropped_rasters", "nut_mask.tif"))
  # Read this in again...
  nut_mask <- raster(here("cropped_rasters", "nut_mask.tif"))
  # Reproject the masked raster to our crs
  nut_reproject <- projectRaster(nut_mask, crs = proj, progress = "text")
  # Save the reprojected layer
  writeRaster(nut_reproject, here("cropped_rasters", "OHI_nutrient_pollution_reproject.tif"))
}

nut_reproject <- raster(here("cropped_rasters", "OHI_nutrient_pollution_reproject.tif"))

# This resamples the raster to 2 degrees, using a bilinear method
if(!file.exists(here("cropped_rasters", "OHI_nut_resample_2dg.tif"))){
  nut_resample <- raster::resample(nut_reproject, x, method = "bilinear", progress = "text")
  nut_resample_mask <- mask_fx(nut_resample)
  writeRaster(nut_resample_mask, here("cropped_rasters", "OHI_nut_resample_2dg.tif"))
}

nut_resample_mask <- raster(here("cropped_rasters", "OHI_nut_resample_2dg.tif"))

plot(nut_resample_mask)
```


### Nitrate from Halley's paper.
This is originally in wgs84.
This is super low resolution (1 degree.)
These just include the mean of 0-10m measurements.

#### DANGER DANGER THIS LOOKS LIKE IT HAS MORE RESOLUTION THAN IT ACTUALLY DOES

```{r, eval = FALSE}
# Let's just use this...
if(!file.exists(here("cropped_rasters", "nitro_mean_2dg_2013.tif"))){
  # Read it in as a raster...
  fall_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Fall.tif"))
  # Crop and mask it to the US EEZ:
  fall_nut_mask <- mask_fx(fall_nut)
  # Repeat for spring, summer, fall
  # SPRING
  # Read it in as a raster...
  spring_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Spring.tif"))
  # Crop and mask it to the US EEZ:
  spring_nut_mask <- mask_fx(spring_nut)
  # SUMMER
  summer_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Summer.tif"))
  # Crop and mask it to the US EEZ:
  summer_nut_mask <- mask_fx(summer_nut)
  # WINTER
  winter_nut <- raster(here("data", "frazier_nut", "avg_nitrate_Winter.tif"))
  # Crop and mask it to the US EEZ:
  winter_nut_mask <- mask_fx(winter_nut)
  
  # Stack these rasters:
  nitro_stack <- raster::stack(fall_nut_mask, spring_nut_mask, summer_nut_mask, winter_nut_mask)
   # Then, find the mean...
  nut_mean <- raster::calc(nitro_stack, fun = mean)
  # Then, resample down to the 0.2 degree resolution
  nitro_mean_resample <- raster::resample(nut_mean, x)
  # Then, recrop and mask:
  nitro_mean_resample_mask <- mask_fx(nitro_mean_resample)
  # Then, write raster
  writeRaster(nitro_mean_resample_mask, here("cropped_rasters", "nitro_mean_2dg_2013.tif"))
}
```

```{r}
nitro_mean <- raster(here("cropped_rasters", "nitro_mean_2dg_2013.tif"))
plot(nitro_mean, nut_resample_mask)
```


### Phosphate from Halley's paper.
This is originally in wgs84.
This is super low resolution (1 degree.)
These just include the mean of 0-10m measurements.

#### DANGER DANGER THIS LOOKS LIKE IT HAS MORE RESOLUTION THAN IT ACTUALLY DOES

```{r, eval = FALSE}
# Let's just use this...
if(!file.exists(here("cropped_rasters", "phos_mean_2dg_2013.tif"))){
  # Read it in as a raster...
  fall_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Fall.tif"))
  # Crop and mask it to the US EEZ:
  fall_phos_mask <- mask_fx(fall_phos)
  # Repeat for spring, summer, fall
  # SPRING
  # Read it in as a raster...
  spring_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Spring.tif"))
  # Crop and mask it to the US EEZ:
  spring_phos_mask <- mask_fx(spring_phos)
  # SUMMER
  summer_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Summer.tif"))
  # Crop and mask it to the US EEZ:
  summer_phos_mask <- mask_fx(summer_phos)
  # WINTER
  winter_phos <- raster(here("data", "frazier_nut", "avg_phosphate_Winter.tif"))
  # Crop and mask it to the US EEZ:
  winter_phos_mask <- mask_fx(winter_phos)
  
  # Stack these rasters:
  phos_stack <- raster::stack(fall_phos_mask, spring_phos_mask, summer_phos_mask, winter_phos_mask)
   # Then, find the mean...
  phos_mean <- raster::calc(phos_stack, fun = mean)
  # Then, resample down to the 0.2 degree resolution
  phos_mean_resample <- raster::resample(phos_mean, x, method = )
  # Then, recrop and mask:
  phos_mean_resample_mask <- mask_fx(phos_mean_resample)
  # Then, write raster
  writeRaster(phos_mean_resample_mask, here("cropped_rasters", "phos_mean_2dg_2013.tif"))
}
```

```{r}
phos_mean <- raster(here("cropped_rasters", "phos_mean_2dg_2013.tif"))
```

```{r}
plot(phos_mean)
plot(US_crop, add = TRUE, col = "grey80")
```

## Okay, there's missing data in these rasters, especially for shoreline values. I'm going to interpolate using an Inverse Distance Weighting function and methods/code from (OHI)[http://ohi-science.org/ohiprep_v2019/globalprep/prs_oa/v2017/create_oa_layer.html].

I need to check and see if this is legit...
```{r}
if(!file.exists(here("cropped_rasters", "phos_int_mask.tif"))){
  inter <- function(file){
    r <- file             # Load in the raster
    # grid <- raster::projectExtent(r, res = 0.2)
    xy <- data.frame(xyFromCell(r, 1:ncell(r))) #create a df of the grid
    v = getValues(r) # and a list of the values..
    tmpdf <- cbind(xy, v)%>%filter(!is.na(v)) # Bind them together
    mg <- gstat(id = "v", formula = v~1, locations = ~x+y, data=tmpdf,
              nmax=7, set=list(idp = 2)) # Create a model (power function = 2)
    z <- interpolate(r, mg, progress='text')
  }
  
  # Run this for both phosphate and nitrate
  phos_inter <- inter(phos_mean)
  nitro_inter <- inter(nitro_mean)
  
  # Mask the values to the US EEZ: 
  phos_int_mask <- mask_fx(phos_inter)
  nitro_int_mask <- mask_fx(nitro_inter)
  
  writeRaster(phos_int_mask, here("cropped_rasters", "phos_int_mask.tif"))
  writeRaster(nitro_int_mask, here("cropped_rasters", "nitro_int_mask.tif"))
}
```


## Bio Oracle Data:

I've found finer resolution data from BioOracle. I'm gonna repeat the same analysis with this:

these are originally 0.083 degree cells
This is already in WGS84

Start with phosphorous:
```{r}
if(!file.exists(here("cropped_rasters", "BO_phos_2dg.tif"))){
  # Load phosphate data in
  bo_phos <- raster(here("data", "BO_phosphate.tif"))
  # Reproject to 2 degree
  bo_phos_2dg <- projectRaster(bo_phos, x, fun = "mean")
  # crop and mask
  bo_phos_mask <- mask_fx(bo_phos_2dg)
  # Save it
  writeRaster(bo_phos_mask, here("cropped_rasters", "BO_phos_2dg.tif"))
}
```

Then nitrogen:
```{r}
if(!file.exists(here("cropped_rasters", "BO_nit_2dg.tif"))){
  # Load phosphate data in
  bo_nit <- raster(here("data", "BO_nitrate.tif"))
  # Reproject to 2 degree
  bo_nit_2dg <- projectRaster(bo_nit, x, fun = "mean")
  # crop and mask
  bo_nit_mask <- mask_fx(bo_nit_2dg)
  # Save it
  writeRaster(bo_nit_mask, here("cropped_rasters", "BO_nit_2dg.tif"), overwrite = TRUE)
}
```

```{r}
bo_nit_mask <- raster(here("cropped_rasters", "BO_nit_2dg.tif"))
bo_phos_mask <- raster(here("cropped_rasters", "BO_phos_2dg.tif"))
```

```{r}
plot(bo_nit_mask)
```
```{r}
plot(bo_phos_mask)
```

## Outfall data:
This data is the total estimated yearly input of either nitrogen or phosphorous from a point source.
This is in mercator projection.
ttl_p__ is the layer name for total pounds of nutrient, for the whole year

```{r}

if(!file.exists(here("cropped_rasters", "outfall_nitrogen.tif"))){
  # Read in nitrogen and phosphorous outfall data:
  outfall_n <- st_read(here("data", "outfall_data"), layer = "outfall_nitrogen") %>% 
    dplyr::select(ttl_p__)
  outfall_p <- st_read(here("data", "outfall_data"), layer = "outfall_phosphorous") %>% 
    dplyr::select(ttl_p__)
  # Reproject these data to WGS84:
  out_n_reproj <- st_transform(outfall_n, proj)
  out_p_reproj <- st_transform(outfall_p, proj)
  # Turn them into rasters:
  out_n_raster <- rasterize(out_n_reproj, x, fun = sum)
  out_p_raster <- rasterize(out_p_reproj, x, fun = sum)
  # Crop and mask them to the US EEZ
  out_n_mask <- mask_fx(out_n_raster)
  out_p_mask <- mask_fx(out_p_raster)
  # Save the cropped rasters:
  writeRaster(out_n_mask, here("cropped_rasters", "outfall_nitrogen.tif"), overwrite = TRUE)
  writeRaster(out_p_mask, here("cropped_rasters", "outfall_phosphorous.tif"), overwrite = TRUE)
}

```


# The following layers are unused in the preliminary analysis and were added later.

## Deposition rates

NOAA's data from 1993?

These are in a 5 degree by 3.75 degree grid...aka pretty terrible. 
also...not projected. But this is the industry standard dataset...
Units are `mg N/m2/yr`

```{r}
if(!file.exists(here("cropped_rasters", "nit_dep_1993.tif"))){
  # Read in
  nit_dep_1993 <- raster(here("data", "anthro_nit_from_noaa", "N-deposition1993.tif"))
  # Set the crs to WGS84
  crs(nit_dep_1993) <- proj
  # Is this the right CRS? I Don't Know!
  plot(nit_dep_1993)
  # To make sure I'm projecting this right, I'm going to first create a raster with the intended resolution and projection...
  dep_raster <- raster(xmn=-180, xmx=180, ymn=-90, ymx=90, res= c(5, 3.75), crs= proj)
  # Then, project the original raster to the correct grid
  nit_1993_b <- projectRaster(nit_dep_1993, dep_raster)
  # Check to make sure this works
  plot(nit_1993_b)
  plot(US_crop, add = TRUE)
  # Then mask it to our AOI.
  nit_1993_crop <- mask_fx(nit_1993_b)
  # And save it:
  writeRaster(nit_1993_crop, here("cropped_rasters", "nit_dep_1993.tif"))
}
```

### Ocean Acidification: 

```{r}

arag <- brick(here("data", "NOAA aragonite saturation statte", "Aragonite.nc"), varname = "Aragonite")

plot(arag)
```

### Seaweed shapefiles

These are calculated seaweed distributions from...Ilan? Ian?
At least from emlab

```{r}
# Read in all the seaweed distributions:
eucheuma <- read_sf(here("data", "distribution-shapefiles"), layer = "Eucheuma")
gracilaria <- read_sf(here("data", "distribution-shapefiles"), layer = "Gracilaria")
kappaphycus <- read_sf(here("data", "distribution-shapefiles"), layer = "Kappaphycus")
macrocystis <- read_sf(here("data", "distribution-shapefiles"), layer = "Macrocystis")
pyropia <- read_sf(here("data", "distribution-shapefiles"), layer = "Pyropia")
saccharina <- read_sf(here("data", "distribution-shapefiles"), layer = "Saccharina")
sargassum <- read_sf(here("data", "distribution-shapefiles"), layer = "Sargassum")
ulva <- read_sf(here("data", "distribution-shapefiles"), layer = "Ulva")
undaria <- read_sf(here("data", "distribution-shapefiles"), layer = "Undaria")
# Combine these all in a data frame:
sp_dist <- rbind(eucheuma, gracilaria, kappaphycus, macrocystis, pyropia, saccharina, sargassum, ulva, undaria)
# Project it to our crs:
sp_dist_tsfm <- st_transform(sp_dist, proj)
# Crop to the US EEZ:
sp_dist_crop <- st_crop(sp_dist, us_eez)
# then mask it to the US EEZ:
sp_dist_int <- st_intersection(sp_dist, us_eez)

sp_dist_int_ak <- st_intersection(sp_dist, ak_eez)

sp_dist_int_hi <- st_intersection
```

